\Large{Research Seminar Software Science} \\ \huge{Simulation of monadic effects in Haskell}

### 2018-05-03 paper.tex
explained the bind operator >>=

    \section{Lists}
    This section shows how the monadic behaviour of lists can be used in our advantage to explicitly demonstrate non-determinism. \\ \\
    In general, lists can be both \textit{deterministic} and \textit{non-deterministic}. For instance, take into consideration a list containing just one element \textbf{[27]}. Clearly, the former list is deterministic since it accommodates a single value and it is apparent that the value equals to 27. On the other hand, a list enclosing more than one element \textbf{[27, 6, 21]} introduces non-deterministic behaviour since it can have many outcomes at the same time. \\ \\
    The notion of \textit{non-determinism} can be decoded smoothly in monadic terms. Below is illustrate how the Monad instance is defined for a list: \\
    \begin{lstlisting}[language=Haskell]
        instance Monad [] where
            return x = [x]
            xs >>= f = concat (map f xs)
    \end{lstlisting} 
    The \textbf{return} wraps a values of type \textit{x} and places it into the special context of a list. In other words, it creates a list that contains only that single value. Afterwards, this not so simple value anymore can interact with other values to produce a non-deterministic nature.   \\ \\
    The bind operator \textbf{($>>$=)} is taking a monadic list, removing its elements from the monadic context and feeding them into a function that takes normal values and returns again monadic list.  In order to remove the confusion, the example below can be used to clarify the complexity of the bind operator: \\
    \begin{lstlisting}[language=Haskell]
       ghci> [3,6,10] >>= \x -> [x, 2*x]
       [3,6,6,12,10,20]
    \end{lstlisting} \\

>>>>

    \section{Lists}
    This section shows how the monadic behaviour of lists can be used in our advantage to explicitly demonstrate non-determinism. \\ \\
    In general, lists can be both \textit{deterministic} and \textit{non-deterministic}. For instance, take into consideration a list containing just one element \textbf{[27]}. Clearly, the former list is deterministic since it accommodates a single value and it is apparent that the value equals to 27. On the other hand, a list enclosing more than one element \textbf{[27, 6, 21]} introduces non-deterministic behaviour since it can have many outcomes at the same time. \\ \\
    The notion of \textit{non-determinism} can be decoded smoothly in monadic terms. Below is illustrate how the Monad instance is defined for a list: \\
    \begin{lstlisting}[language=Haskell]
        instance Monad [] where
            return x = [x]
            xs >>= f = concat (map f xs)
    \end{lstlisting} 
    The \textbf{return} wraps a values of type \textit{x} and places it into the special context of a list. In other words, it creates a list that contains only that single value. Afterwards, this not so simple value anymore can interact with other values to produce a non-deterministic nature.   \\ \\
    The bind operator \textbf{($>>$=)} is taking a monadic list, removing its elements from the monadic context and feeding them into a function that takes normal values and returns again monadic list.  In order to remove the confusion, the example below can be used to clarify the complexity of the bind operator: \\
    \begin{lstlisting}[language=Haskell]
       ghci> [3, 6, 10] >>= \x -> [x, 2*x]
       [3, 6, 6, 12, 10, 20]
    \end{lstlisting} \\
    Therefore, the purpose of the bind operator is to facilitate the non-deterministic attitude of lists. \textbf{[3, 6, 10]} is a non-deterministic list that is being feed into a function that itself returns a non-deterministic list. Hence, the final result will be a non-deterministic list containing all the possible combinations between the elements of the list \textbf{[3, 6, 10]} and the passing of these values to the function \textbf{\textbackslash x $\rightarrow$ [x, 2*x]}. The job of the function is straightforward. The function generates two outcomes for each element of the input list, the original value and the value multiplied by 2.
